Sensor Clusterization in D-optimal Design in Infinite Dimensional Bayesian Linear Inverse Problems

### 2024-09-08 latex/intro.tex
Done

\begin{enumerate}
\item \label{q:generic} \textbf{Is measurement clusterization a
  generic phenomenon?}
  %\subsection{An answer for Question \ref{q:generic}: Genericity of measurement clusterization}
  %% Computer implementation of Lemma \ref{thm:char} for the inverse
  %% problem outlined in Section \ref{section:how} also generates \(\obs\)
  %% as a solution to the D-optimal design problem.  Furthermore,
  We give two complementing answers to this question. First, from a
  theoretical perspective, we show that clusterization mainly depends
  on how quickly the eigenvalues of the prior covariance in
  observation space decay.
  %% This decay depends mostly on how ill-posed the problem at hand is
  %% and does not depend much on the prior
  See Section \ref{section:vanishing} --- particularly, Theorem
  \ref{thm:char} and the discussion following it. Furhtermore, in
  Section \ref{subsec:lemma_sims} we show results of numerical
  experiments, where simulations of our model give rise to D-optimal
  designs that exhibit clusterization with high probability. Thus,
  given the genericity of our model, we expect measurement
  clusterization to be a generic and ubiquitous phenomenon.

@@

%% Linearization
Another potential cause of clusterization is linearization. While in a
real applications the model is not necessarily linear, some authors
consider a linearized version of their model when seeking optimal
designs \cite{fedorov1996, neitzel2019sparse}. Our analysis then shows
that the linearity of the forward problem is also an important
ingredient in giving rise to clustered desgins. Therefore, our advice
to practitioners is to avoid linearization, and find D-optimal designs
via other methods, e.g.~the sampling method of \cite{ryan2003}.

%% Avoid
Given the above discussion, it is our belief that the methods
suggested by other authors to avoid clusterization are merely
overlooking the problem. We do not view clustered designs as
inherently bad on their own. Rather, we suggest that if a clustered
design arises, a practitioner should revisit their modeling
assumptions --- specifically, their choice of prior and model
linearization. If the practitioner is confident in their assumptions,
then clustered designs should be avoided only to the extent
necessitated by the physical measuring apparatus.

%% Convergence
Clusterization in D-optimal designs raises the question of whether
D-optimal designs should be pursued at all. While other authors have
established convergence results for decaying measurement error
\cite{knapik2011}, space-filling measurements \cite{teckentrup2020}
and randomly sampled designs \cite{nickl2023}, their results do not
hold for D-optimal designs. %% Unfortunately, D-optimal designs ---
%% clustered designs in particular --- do not fall into any of these
%% categories.

Luckily, we can expect convergence from D-optimal designs (including
clustered designs). Convergence holds in the following sense: the
posterior uncertainty ellipsoid will contract to zero along each
one of its eigenvectors. See Section \ref{subsub:convergence} for a
precise statement and a proof based on Theorem \ref{thm:char} --- the
main theorem of this manuscript.

>>>>

\begin{enumerate}
\item \label{q:generic} \textbf{Is measurement clusterization a
  generic phenomenon?}
  %\subsection{An answer for Question \ref{q:generic}: Genericity of measurement clusterization}
  %% Computer implementation of Lemma \ref{thm:char} for the inverse
  %% problem outlined in Section \ref{section:how} also generates \(\obs\)
  %% as a solution to the D-optimal design problem.  Furthermore,
  We give two complementing answers to this question. First, from a
  theoretical perspective, we show that clusterization mainly depends
  on how quickly the eigenvalues of the prior covariance in
  observation space decay.
  %% This decay depends mostly on how ill-posed the problem at hand is
  %% and does not depend much on the prior
  See Section \ref{section:vanishing} --- particularly, Theorem
  \ref{thm:char} and the discussion following it. Furthermore, in
  Section \ref{subsec:lemma_sims} we show results of numerical
  experiments, where simulations of our model give rise to D-optimal
  designs that exhibit clusterization with high probability. Thus,
  given the genericity of our model, we expect measurement
  clusterization to be a generic and ubiquitous phenomenon.

@@

%% Linearization
Another potential cause of clusterization is linearization. While in a
real applications the model is not necessarily linear, some authors
consider a linearized version of their model when seeking optimal
designs \cite{fedorov1996, neitzel2019sparse}. Our analysis then shows
that the linearity of the forward problem is also an important
ingredient in giving rise to clustered designs. Therefore, our advice
to practitioners is to avoid linearization, and find D-optimal designs
via other methods, e.g.~the sampling method of \cite{ryan2003}.

%% Avoid
Given the above discussion, it is our belief that the methods
suggested by other authors to avoid clusterization are merely
overlooking the problem. We do not view clustered designs as
inherently bad on their own. Rather, we suggest that if a clustered
design arises, a practitioner should revisit their modeling choices
--- specifically, their choice of prior and model linearization. If
the practitioner is confident in their analysis, then clustered
designs should be avoided only to the extent necessitated by the
physical measuring apparatus.

%% Convergence
Clusterization in D-optimal designs raises the question of whether
D-optimal designs should be pursued at all. While other authors have
established convergence results for decaying measurement error
\cite{knapik2011}, space-filling measurements \cite{teckentrup2020}
and randomly sampled designs \cite{nickl2023}, their results do not
hold for D-optimal designs. Luckily, we can expect convergence from
D-optimal designs (including clustered designs) in the following
sense: the posterior uncertainty ellipsoid will contract to zero along
each one of its eigenvectors. See Section \ref{subsub:convergence} for
a precise statement and a proof based on Theorem \ref{thm:char} ---
the main theorem of this manuscript.

### 2024-09-08 revision/refs.tex
Done

I appreciate the many thoughtful insightful comments by all of
you. These comments have helped me considerably improve the exposition
of my paper and added new insights as well. I sincerely appreciate the
option to revise and resubmit this manuscript. Please see below my
answers to the comments made by the editor in chief (Section
\ref{eic}), associate editor (Section \ref{ae}) and three referees
(Sections \ref{ref1}, \ref{ref2} and \ref{ref3}).

@@

\AR I have considerably extended the introduction. See revised
manuscript, diff file and detailed answers to referees' comments
below.

@@

\AR I added a section on the implications. The main takeaways are: (1)
avoid linearization and Gaussian priors, and (2) clusterization can
inform you of the number of relevant d.o.f in the problem.

@@

\AR I have added considerably to the introduction. I hope you will
find the transfer to the maths smoother now.

@@

\AR Again, I appreciate the time and effort invested by the referees
and editorial team. Thanks for the opportunity to revise and improve this manuscript.

>>>>

I appreciate the many insightful comments by the refrees. These
comments have helped me considerably improve the exposition of the
manuscript and added new insights as well. I sincerely appreciate the
possibility to revise and resubmit this manuscript. Please see below
my answers to the comments made by the editor in chief (Section
\ref{eic}), associate editor (Section \ref{ae}) and three referees
(Sections \ref{ref1}, \ref{ref2} and \ref{ref3}).

@@

\AR I have considerably extended the Introduction. See revised
manuscript, diff file and detailed answers to referees' comments
below.

@@

\AR I added a section discussing implications. The main takeaways are:
(1) avoid linearization and Gaussian priors, and (2) clusterization
can inform you of the number of relevant d.o.f in the problem.

@@

\AR I have added considerably to the Introduction. I hope you will
find the transfer to the maths smoother now.

@@

\AR Again, I appreciate the time and effort invested by the referees
and editorial team. Thank you for the opportunity to revise and
improve this manuscript.

### 2024-09-08 revision/rev1.tex
Typos
Done

\section{Reviewer 1}\label{ref1}
\RC This paper explains a common practical issue: the cause of
clusterization in Bayesian D-optimal designs in infinite dimensions
and why adding a correlated measurement process decluster the design
points. Its theoretical results are valuable to the field of inverse
problems/uncertainty quantification and the conclusions should be
useful to practitioners in this field.

@@

\begin{quote} % Avoid
Given the above discussion, it is our belief that the methods
suggested by other authors to avoid clusterization are merely
overlooking the problem. We do not view clustered designs as
inherently bad on their own. Rather, we suggest that if a clustered
design arises, a practitioner should revisit their modeling
assumptions --- specifically, their choice of prior and model
linearization. If the practitioner is confident in their assumptions,
then clustered designs should be avoided only to the extent
necessitated by the physical measuring apparatus.
\end{quote}

>>>>

\section{Referee 1}\label{ref1}
\RC This paper explains a common practical issue: the cause of
clusterization in Bayesian D-optimal designs in infinite dimensions
and why adding a correlated measurement process decluster the design
points. Its theoretical results are valuable to the field of inverse
problems/uncertainty quantification and the conclusions should be
useful to practitioners in this field.

@@

\begin{quote} % Avoid
Given the above discussion, it is our belief that the methods
suggested by other authors to avoid clusterization are merely
overlooking the problem. We do not view clustered designs as
inherently bad on their own. Rather, we suggest that if a clustered
design arises, a practitioner should revisit their modeling choices
--- specifically, their choice of prior and model linearization. If
the practitioner is confident in their analysis, then clustered
designs should be avoided only to the extent necessitated by the
physical measuring apparatus.
\end{quote}

### 2024-09-08 revision/rev2.tex
Typos
Done

\section{Reviewer 2}\label{ref2}
\subsection{Measurement Clusterization vs Repetition}
\RC I will start by saying measurement clusterisation is not
well-defined in this paper. On page 1, it is referred to as two sets
of measurements that are almost indistinguishable from one
another. However, the setup for Corollary 11 implicitly defines
measurement clusterisation as repeated measurements. The latter
definition matches the references of \cite{fedorov1996} and
\cite{nyberg2012}. So I will assume for the rest of this review that,
measurement clusterisation refers to repeated measurements.

@@

\begin{quote} %% Replication
  Clusterization should not be confused with replication. Replication
  requires that the experimentalist executes multiple trials under
  circumstances that are \emph{nominally identical} \cite[Section
    1.2.4]{morris2011}. Replication is commonly viewed as a beneficial
  and even necessary aspect of optimal experimental design
  \cite{fisher1949design, morris2011,
    schafer2001replication}. Similarly to a design implementing
  replication, a clustered design reduces the signal-to-noise ratio of
  the repeated measurements \cite{telford2007brief}. The difference is
  that a clustered design takes repeated measurements at the expense
  of other quantities not measured at all.
  \newline
For example, \cite{fisher1949design}, suggested repeating his famous
milk and tea experiment in order "to be able to demonstrate the
predominance of correct classifications in spite of occasional
errors". Unfortunately, in the experiments we consider, replication is
impossible. For example, in the MRI problem, we cannot generate an
individual nominally identical to the one we wish to scan.
\newline
To further illustrate the difference between replication and
clusterization, consider an experiment measuring the effect of
rainfall on grass growth \cite{fay2000rainfall}. The experiment
involved four rainfall manipulation "treatments" (i.e.~simulating
different timing and quantity of rainfall), each replicated three
times over a different plot of land. Indeed, it seems reasonable for
researchers to replicate the phenomenon they are trying to study. A
clustered design in such an experiment would imply the researchers
should take a repeated measurement \emph{on the same plot}, at the
expense of measuring grass growth in other plots. To conclude our
short discussion on replication vs.~clusterization: these are
fundamentally different concepts and even though replication is quite
intuitive, it is inapplicable to the inverse problems we consider in
this manuscript.
\end{quote}

@@

\AR I agree, but what you are describing is replication. See my
comment on replication vs.~clusterization above.

@@

\AR I really appreciate this comment. I added a section highlighting
the implications of this study. Since it is too long to reproduce
here, see Section \ref{subsub:implications} in the revised manuscript.

@@

\AR Thanks for this comment. I considerably expanded the
  introduction and delegated many proofs to the Supplementary.

@@

\AR To a certain extent I agree: the impossibility of taking identical
measurements is indeed natural. However, this is not encoded in the
mathematical formulation of the problem. One can then take one of two
approaches: either accept what the mathematics suggest, even if it
counterintuitive, or find ways to circumvent the mathematics because
the result seems non-intuitive. I expanded this discussion in the
revised manuscript. See below and the discussion preceding it in
Section \ref{subsub:implications} of the revised manuscript.

>>>>

\section{Referee 2}\label{ref2}
\subsection{Measurement Clusterization vs Repetition}
\RC I will start by saying measurement clusterisation is not
well-defined in this paper. On page 1, it is referred to as two sets
of measurements that are almost indistinguishable from one
another. However, the setup for Corollary 11 implicitly defines
measurement clusterisation as repeated measurements. The latter
definition matches the references of \cite{fedorov1996} and
\cite{nyberg2012}. So I will assume for the rest of this review that,
measurement clusterisation refers to repeated measurements.

@@

\begin{quote} %% Replication
  Clusterization should not be confused with replication. Replication
requires that the experimentalist executes multiple trials under
circumstances that are \emph{nominally identical} \cite[Section
  1.2.4]{morris2011}. Replication is commonly viewed as a beneficial
and even necessary aspect of optimal experimental design
\cite{fisher1949design, morris2011, schafer2001replication}.
For example, \cite{fisher1949design}, suggested repeating his famous
milk and tea experiment in order "to be able to demonstrate the
predominance of correct classifications in spite of occasional
errors". Unfortunately, in the experiments we consider, replication is
impossible. For example, in the MRI problem, we cannot generate an
individual nominally identical to the one we wish to scan.
\newline
Similarly to a design implementing replication, a clustered design
reduces the signal-to-noise ratio of the repeated measurements
\cite{telford2007brief}. The difference is that a clustered design
takes repeated measurements at the expense of other quantities not
measured at all. For example, consider an experiment measuring the
effect of rainfall on grass growth \cite{fay2000rainfall}. The
experiment involved four rainfall manipulation "treatments"
(i.e.~simulating different timing and quantity of rainfall), each
replicated three times over different plots of land. Indeed, it seems
reasonable for researchers to replicate the phenomenon they are trying
to study. A clustered design in such an experiment would imply the
researchers should take a repeated measurement \emph{on the same
plot}, at the expense of measuring grass growth in other plots!
\newline
To conclude our short discussion on replication vs.~clusterization:
these are fundamentally different concepts and even though replication
is quite intuitive, it is inapplicable to the inverse problems we
consider in this manuscript.
\end{quote}

@@

\AR I agree, but I believe what you are describing is replication. See
my comment on replication vs.~clusterization above.

@@

\AR I really appreciate this comment. I added a section highlighting
the implications of this study. Since it is too long to reproduce
here, see Section \ref{subsub:implications} in the revised
manuscript. I will just mention that one important takeaway is that
practitioners should avoid linearization and Gaussian priors when
seeking D-optimal designs.

@@

\AR Thanks for this comment. I considerably expanded the
  Introduction and delegated many proofs to the Supplementary.

@@

\AR To a certain extent I agree: the impossibility of taking identical
measurements is indeed natural. However, this is not encoded in the
mathematical formulation of the problem. One can then take one of two
approaches: either accept what the mathematics suggest, even if it
counterintuitive, or find ways to circumvent the mathematics because
the result seems non-intuitive. I expanded this discussion in the
revised manuscript. See paragraph below and the discussion preceding
it in Section \ref{subsub:implications} of the revised manuscript.
